A Bias-Variance Decomposition of Information Transfer in Multi-Task Learning

### 2020-08-27 arxiv/transfer.tex

\begin{lemma}[Variance asymptotics]\label{lem_cov_shift}
	%Let $X_i\in\real^{n_i\times p}$ be a random matrix that contains i.i.d. row vectors with mean $0$ and variance $\Sigma_i\in\real^{p\times p}$, for $i = 1, 2$.
	%Suppose $X_1=Z_1\Sigma_1^{1/2}\in \R^{n_1\times p}$ and $X_2=Z_2\Sigma_2^{1/2}\in \R^{n_2\times p}$ satisfy Assumption \ref{assm_secA1} with $\rho_1:=n_1/p>1$ and $\rho_2:=n_2/p>1$ being fixed constants.
	%Denote by $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ and
	Let $\Sigma \in \real^{p \times p}$ be any fixed and deterministic matrix.
	In the setting of Theorem \ref{thm_main_informal},
	with probability $1-\oo(1)$ over the randomness of $X_1$ and $X_2$, we have that %for any constant $\e>0$,
	%When $n_1 = c_1 p$ and $n_2 = c_2 p$, we have that with high probability over the randomness of $X_1$ and $X_2$, the following equation holds
	\begin{align}\label{lem_cov_shift_eq}
		\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} \Sigma} = \frac{1}{n_1 + n_2}\cdot \bigtr{ (a_1 \Sigma_1 + a_2\Sigma_2)^{-1} \Sigma} + o(\norm{\Sigma}), %\bigo{\|\| p^{-1/2+\epsilon}}
	\end{align}
	where $(a_1, a_2)$ is the solution of the following deterministic equations:
	\begin{align}
		a_1 + a_2 = 1- \frac{1}{\rho_1 + \rho_2},\quad a_1 + \frac1{\rho_1 + \rho_2}\cdot \bigbrace{\frac{1}{p}\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{\rho_1}{\rho_1 + \rho_2}. \label{eq_a12extra}
	\end{align}
\end{lemma}

@@

\begin{lemma}[Bias asymptotics]\label{lem_cov_derivative}
In the setting of Theorem \ref{thm_main_informal}, let $w \in \R^p$ be any vector that is independent of $X_1$ and $X_2$.
With high probability, we have that
\begin{equation}\label{lem_cov_derv_eq}
\begin{split}
\bignorm{\Sigma_2^{1/2} \bigbrace{\frac{X_1^{\top}X_1 + X_2^{\top}X_2}{n_1 + n_2}}^{-1} w}^2
= w^{\top} \Sigma_2^{-1/2}  \frac{a_4 {M}^{\top}{M} + (1 + a_3)\id}{(a_1 {M}^{\top}{M}+a_2\id)^2} \Sigma_2^{-1/2} w + o(\|w\|^2),
\end{split}
\end{equation}
where $a_{3}$ and $a_4$ satisfy the following system of linear equations:
\begin{gather}\label{eq_a34extra}
		\left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_3 - b_1 \cdot  a_4
		= b_0, \quad \left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_4 -  b_1 \cdot  a_3 = b_1 .
%		\left(\frac{n_1}{\hat a_1^2} -  \sum_{i=1}^p \frac{\hat \lambda_i^4   }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_4 -\left(\sum_{i=1}^p \frac{\hat \lambda_i^2  }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_3
%		= \sum_{i=1}^p \frac{\hat \lambda_i^2 }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }. \label{eq_a4}
	\end{gather}
Here $b_0$, $b_1$ and $b_2$ are defined as
$$ b_k:= \frac1{p}\sum_{i=1}^p \frac{\lambda_i^{2k}}{ ( a_2 + \lambda_i^2 a_1)^2  },\quad  \text{for } k=0,1,2.$$
\end{lemma}
To apply the above result to equation \eqref{eq_bias_2task}, we first approximate $X_1^{\top}X_1$ by $\Sigma_1$, and then use Lemma \ref{lem_cov_derivative} with $w = \Sigma_1 (\beta_1 - \frac{W_1}{W_2} \beta_2)$.
The approximation of $X_1^{\top}X_1$ by $\Sigma_1$ incurs a generalization error that scales down with $\rho_1$, which is the $\delta / \sqrt{\rho_1}$ term in Theorem \ref{thm_main_informal} more precisely.

>>>>

\begin{lemma}[Variance asymptotics]\label{lem_cov_shift}
	%Let $X_i\in\real^{n_i\times p}$ be a random matrix that contains i.i.d. row vectors with mean $0$ and variance $\Sigma_i\in\real^{p\times p}$, for $i = 1, 2$.
	%Suppose $X_1=Z_1\Sigma_1^{1/2}\in \R^{n_1\times p}$ and $X_2=Z_2\Sigma_2^{1/2}\in \R^{n_2\times p}$ satisfy Assumption \ref{assm_secA1} with $\rho_1:=n_1/p>1$ and $\rho_2:=n_2/p>1$ being fixed constants.
	%Denote by $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ and
	Let $\Sigma \in \real^{p \times p}$ be any fixed and deterministic matrix.
	In the setting of Theorem \ref{thm_main_informal},
	with probability $1-\oo(1)$ over the randomness of $X_1$ and $X_2$, we have that %for any constant $\e>0$,
	%When $n_1 = c_1 p$ and $n_2 = c_2 p$, we have that with high probability over the randomness of $X_1$ and $X_2$, the following equation holds
	\begin{align}\label{lem_cov_shift_eq}
		\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} \Sigma} = \frac{1}{n_1 + n_2}\cdot \bigtr{ (a_1 \Sigma_1 + a_2\Sigma_2)^{-1} \Sigma} + o(\norm{\Sigma}). %\bigo{\|\| p^{-1/2+\epsilon}}
	\end{align}
	In the above equation, $a_1$ and $a_2$ are the solutions of the following two equations that only depend on the sample sizes $\rho_1, \rho_2$, and the singluar values of the covariate shift matrix $M$:
	\begin{align}
		a_1 + a_2 = 1- \frac{1}{\rho_1 + \rho_2},\quad a_1 + \frac1{\rho_1 + \rho_2}\cdot \bigbrace{\frac{1}{p}\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{\rho_1}{\rho_1 + \rho_2}. \label{eq_a12extra}
	\end{align}
\end{lemma}

@@

\begin{lemma}[Bias asymptotics]\label{lem_cov_derivative}
In the setting of Theorem \ref{thm_main_informal}, let $w \in \R^p$ be any vector that is independent of $X_1$ and $X_2$.
With high probability, we have that
\begin{equation}\label{lem_cov_derv_eq}
\begin{split}
\bignorm{\Sigma_2^{1/2} \bigbrace{\frac{X_1^{\top}X_1 + X_2^{\top}X_2}{n_1 + n_2}}^{-1} w}^2
= w^{\top} \Sigma_2^{-1/2}  \frac{a_4 {M}^{\top}{M} + (1 + a_3)\id}{(a_1 {M}^{\top}{M}+a_2\id)^2} \Sigma_2^{-1/2} w + o(\|w\|^2),
\end{split}
\end{equation}
where $a_{3}$ and $a_4$ are the solutions of the following two equations:
\begin{gather}\label{eq_a34extra}
		\left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_3 - b_1 \cdot  a_4
		= b_0, \quad \left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_4 -  b_1 \cdot  a_3 = b_1 .
%		\left(\frac{n_1}{\hat a_1^2} -  \sum_{i=1}^p \frac{\hat \lambda_i^4   }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_4 -\left(\sum_{i=1}^p \frac{\hat \lambda_i^2  }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_3
%		= \sum_{i=1}^p \frac{\hat \lambda_i^2 }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }. \label{eq_a4}
	\end{gather}
In the above equation, the definition of $b_0$, $b_1$ and $b_2$ is
$$ b_k:= \frac1{p}\sum_{i=1}^p \frac{\lambda_i^{2k}}{ (\lambda_i^2 a_1 + a_2)^2  },\quad  \text{for } k=0,1,2.$$
\end{lemma}
To apply the above result to equation \eqref{eq_bias_2task}, we first approximate $X_1^{\top}X_1$ by $\Sigma_1$, and then use Lemma \ref{lem_cov_derivative} with $w = \Sigma_1 (\beta_1 - \frac{W_1}{W_2} \beta_2)$.
The approximation of $X_1^{\top}X_1$ by $\Sigma_1$ incurs a generalization error that scales down with $\rho_1$, which is the $\delta / \sqrt{\rho_1}$ term in Theorem \ref{thm_main_informal} more precisely.
